---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

<!-- badges: start -->
[![rcc](https://github.com/krlmlr/mlfit/workflows/rcc/badge.svg)](https://github.com/krlmlr/mlfit/actions)
[![codecov.io](https://codecov.io/github/krlmlr/mlfit/coverage.svg?branch=master)](https://codecov.io/github/krlmlr/mlfit?branch=master)
[![Codecov test coverage](https://codecov.io/gh/krlmlr/mlfit/branch/master/graph/badge.svg)](https://codecov.io/gh/krlmlr/mlfit?branch=master)
<!-- badges: end -->

Implementation of algorithms that extend Iterative Proportional Fitting (IPF) to nested structures.

The IPF algorithm operates on count data. This package offers implementations for several algorithms that extend this to nested structures: "parent" and "child" items for both of which constraints can be provided.

## Example

Here is a multi-level fitting example with a reference sample and two control tables. The reference sample represents a sample of a population. The first control table, `individual_control`, is for individual level and the second table, `group_control`, is for group level. Each row of the reference sample represents an individual, where `HHNR` is their group ID and `PNR` is their individual ID, `APER` and `WKSTAT` are individial-level charateristics, and `CAR` is the only household characteristic of the sample population. The 'N' columns in both control tables denote how many units of individuals or groups belong to each category.

```{r}
library(mlfit)
library(tibble)

reference_sample <- tibble::tribble(
  ~HHNR, ~PNR, ~APER, ~CAR, ~WKSTAT,
     1L,   1L,    3L,  "0",     "1",
     1L,   2L,    3L,  "0",     "2",
     1L,   3L,    3L,  "0",     "3",
     2L,   4L,    2L,  "0",     "1",
     2L,   5L,    2L,  "0",     "3",
     3L,   6L,    3L,  "0",     "1",
     3L,   7L,    3L,  "0",     "1",
     3L,   8L,    3L,  "0",     "2",
     4L,   9L,    3L,  "1",     "1",
     4L,  10L,    3L,  "1",     "3",
     4L,  11L,    3L,  "1",     "3",
     5L,  12L,    3L,  "1",     "2",
     5L,  13L,    3L,  "1",     "2",
     5L,  14L,    3L,  "1",     "3",
     6L,  15L,    2L,  "1",     "1",
     6L,  16L,    2L,  "1",     "2",
     7L,  17L,    5L,  "1",     "1",
     7L,  18L,    5L,  "1",     "1",
     7L,  19L,    5L,  "1",     "2",
     7L,  20L,    5L,  "1",     "3",
     7L,  21L,    5L,  "1",     "3",
     8L,  22L,    2L,  "1",     "1",
     8L,  23L,    2L,  "1",     "2"
  )

individual_control <- tibble::tribble(
  ~WKSTAT,   ~N,
      "1",  91L,
      "2",  65L,
      "3", 104L
  )

group_control <- tibble::tribble(
  ~CAR,  ~N,
   "0", 35L,
   "1", 65L
  )
```

First we need to create a `ml_problem` object which defines our multi-level fitting problem. `special_field_names()` is useful for the `field_names` argument to `ml_problem()`, this is where we need to specific the names of the ID columns in our reference sample and the count column in the control tables.

```{r}
fitting_problem <- ml_problem(
  ref_sample = reference_sample, 
  controls = list(
    individual = list(individual_control),
    group = list(group_control)
  ),
  field_names = special_field_names(
    groupId = "HHNR", 
    individualId = "PNR", 
    count = "N"
  )
)
```

You can use one of the `ml_fit_*()` functions to calibrate your fitting problem, or you can use `ml_fit(ml_problem, algorithm = "<your-selected-algorithm>")`.

```{r}
fit <- ml_fit(ml_problem = fitting_problem, algorithm = "ipu")
fit
```

`mlfit` also provides a function that helps to replicate the reference sample based on the fitted/calibrated weights. See `?ml_replicate` to find out which integerisation algorithms are available.

```{r}
syn_pop <- ml_replicate(fit, algorithm = "trs")
syn_pop
```

## Powered by

- [`grake`](http://krlmlr.github.io/grake): A reimplementation of generalized raking ([Deville and Särndal, 1992](http://amstat.tandfonline.com/doi/abs/10.1080/01621459.1992.10475217); [Deville, Särndal and Sautory, 1993](http://www.tandfonline.com/doi/abs/10.1080/01621459.1993.10476369))


## Related work

- [`wrswoR`](http://krlmlr.github.io/wrswoR): An implementation of fast weighted random sampling without replacement ([Efraimidis and Spirakis, 2006](http://www.sciencedirect.com/science/article/pii/S002001900500298X))
- [`mangow`](http://krlmlr.github.io/mangow): Embed the Gower distance metric in L1
- [`RANN.L1`](https://github.com/jefferis/RANN/tree/master-L1#readme): k-nearest neighbors using the L1 metric


## Installation

The package is not currently on CRAN, but can be installed from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("krlmlr/mlfit")
```

### Where is `MultiLeveLIPF`?

From version `0.4.0` onwards the package is now to be known as `mlfit`. If you would like to install any version that is older than `0.4.0` please use:

``` r
# See https://github.com/krlmlr/mlfit/releases for the releases that are available
# To install a certain branch or commit or tag, append it to the repo name, after an @:
devtools::install_github("krlmlr/MultiLevelIPF@v0.3-7")
```